Automated Dynamic Slice Backsolving

### 2014-03-21 docs/paper/fsuthesis/thesis-template/paper.tex
corrected latex for abstract, some of contributions

%\begin{acknowledgments}
%\end{acknowledgments}

@@

%\begin{abstract}
%This is my abstract.
%\end{abstract}

@@

\chapter{Abstract}

@@

@@

In this paper, we present our analysis of existing techniques along with our own contributions and conclusions. We begin by exploring related work in Chapter 3, which sets the background and describes current fuzzers' techniques and limitations. In chapter 4, we describe our contributions, in the form of what is, to the best of our knowledge, a new technique for black box binary fuzzing, a proof of concept fuzzer which implements the technique, and testing results. Lastly, we describe possible future work and conclusions from the experience in Chapter 5.

@@

% Example for stubble
%\begin{lstlisting}[style=asmstyle]
%call read					;instrument read syscall
%mov eax, [ebp - val]		;taint propagates from val to eax reg
%inc eax						;transform eax
%mul eax, 5					;transform eax
%cmp eax, 42					;cmp and jmp based on tainted reg value
%jnz 0x8BADC0DE
%\end{lstlisting}

%If we already had a test case where the value read onto the stack, after the transformations, did not equal 42, then the cmp instruction would not set the zero flag, and we would jump to the memory location 0xBADC0DE. With dynamic taint analysis in place, we see taint introduced with the read syscall, then moved into the eax register (which is subsequently tainted). After some number of transformations, perhaps zero, to the value in the eax register, we see a compare instruction, which checks eax's value against an immediate value. Then we can keep track of specific taint sources, and can record any movements or transformations, we have all the necessary information to "solve" the branch. Once we reach the tainted branch, we know that user input can affect whether we jump or not, and consequently we can backsolve to find out how to affect the branch instruction. In particular, we know the conditional jump instruction jnz depends on the Z flag in the EFLAGS register. We record whatever instruction last modified each flag in EFLAGS, so we handle the compare instruction next. Because we have kept track of taint propagation, we know we can affect the value of eax, but not 42, since it is an immediate value, so we know we need to make the value in eax equal 42. We follow the instructions that have modified the value of eax and use any appropriate operands involved with the inverse operation of the one originally performed on the value to reverse the transformations to eax's value. Finally, we notice eax originally received taint when a value was moved from memory, which was read into memory using the read syscall.

@@

@@

Stubble architecture

@@

@@

% (segfault? EIP? add vulnerability detection alg?). We have observed .... (insert graphs with password lengths, etc.)

@@

%We are currently instrumenting dummy programs which use a subset of all possible comparisons and arithmetic operations on tainted variables. If results appear promising we can try to make the fuzzer as complete as possible and begin running it on real-world applications. Achieving a noticeable speedup over modern symbolic execution would be the goal.

>>>>

\begin{acknowledgments}
SFS
% Dr. Wang
% Bret Whissel
\end{acknowledgments}

@@

@@

\begin{abstract}

@@

\end{abstract}

@@

In this paper, we present our analysis of existing techniques along with our own contributions and conclusions. We begin by exploring related work in Chapter 2, which sets the background and describes current fuzzers' techniques and limitations. In chapter 3, we describe our contributions, in the form of what is, to the best of our knowledge, a new technique for black box binary fuzzing, a proof of concept fuzzer which implements the technique, and testing results. Lastly, we describe possible future work and conclusions from the experience in Chapter 4.

@@

@@

We have developed a technique which, to the best of our knowledge, provides similar code coverage to symbolic execution by using dynamic taint analysis to record relevant program slices and drive test case mutation. As a program runs, the introduction and propagation of taint is recorded. Should taint flow to a sink of interest, the recorded slice is backsolved to find new byte values which will hopefully lead future program runs down new paths. Taint is generally introduced to the program by user input. This can come in the form of user-provided files, command line arguments, or stdin. When user input is detected, it is marked as tainted at the byte level. This means that if user input is conceptualized as a stream of bytes $b_0,b_1,...,b_n$, then the first byte is considered separate from the separate byte, and each is individually followed throughout program execution. This level of granularity introduces significantly more overhead than binary tainting, which marks bytes as either tainted or not tainted, but provides far greater detail into which bytes of user input affect which parts of a program.

Once user input is captured and marked tainted, any memory accesses, loads and stores, and transformations of tainted bytes are recorded as they are encountered. This creates a slice of moves and alterations to the bytes, which can be followed like a trail of breadcrumbs back to the original position of the byte. This continues as the program executes until a conditional branch is reached. When a conditional branch instruction is reached, it is either taken or not taken, based upon the value of some comparison. If this branch depends upon tainted values, which have descended in one way or another from user input, then there is a strong possibility that the "road not taken" could be explored if only user input looked appropriate. In the case where any transformations to user input are reversible, we can backsolve these transformations and, based upon the comparison, modify the byte value so that it leads to the complement of the previous branch condition. Not only this, we can identify which byte $b_0,b_1,...,b_n$ to modify, since we have followed each byte from their introduction. Using this information, we can mutate the previous test case to create a new test which will result in very similar program behavior, with the exception of branching where before it did not branch, and vice versa.

INSERT COOL FIGURE

The idea for this technique is raised in [A], which suggests using dynamic taint analysis to identify "hot bytes" of user input, which affect important things in a program, and also to generate backward slices [AA]. Hot bytes become the target of fuzzing, with backward slices providing valuable context to analysts. Our contribution is the realization of this idea for a subset of cases, and the automation of solving these slices. Our hope is that by solving these slices automatically, we can identify not only the exact location but also the exact value of hot bytes.

we ignore ....loops, more TODO

Our methodology focuses on following taint specifically up to tainted branches, to direct as much attention as possible to finding new paths to explore within the binary. This is of particular importance when source code is not available for analysis, which is common in vulnerability research. Compiled code can be disassembled or debugged, but compilation is a lossy process. Even industry de-compilers like Hex Rays cannot restore comments, since these are not preserved, and they cannot guarantee to reproduce the exact higher level language semantics, given only assembly language. Thus, our implementation relies on Intel Pin to dynamically disassemble compiled binaries.

Consider the following example in Intel x86 assembly:

% x86 example
\begin{lstlisting}[style=asmstyle]
call read					;instrument read syscall
mov eax, [ebp - val]		;taint propagates from val to eax reg
inc eax						;transform eax
mul eax, 5					;transform eax
cmp eax, 42					;cmp and jmp based on tainted reg value
jnz 0x8BADC0DE
\end{lstlisting}

If we already had a test case where the value read onto the stack, after the transformations, did not equal 42, then the cmp instruction would not set the zero flag, and we would jump to the memory location 0xBADC0DE. With dynamic taint analysis in place, we see taint introduced with the read syscall, then moved into the eax register (which is subsequently tainted). After an arbitrary number of transformations, perhaps zero, to the value in the eax register, comes a compare instruction, which checks eax's value against an immediate value. Then we can keep track of specific taint sources, and can record any movements or transformations, we have all the necessary information to "solve" the branch. Once we reach the tainted branch, we know that user input can affect whether we jump or not, and consequently we can backsolve to find out how to affect the branch instruction. In particular, we know the conditional jump instruction jnz depends on the Z flag in the EFLAGS register. We record whatever instruction last modified each flag in EFLAGS, so we handle the compare instruction next. Because we have kept track of taint propagation, we know we can affect the value of eax, but not 42, since it is an immediate value, so we know we need to make the value in eax equal 42. We follow the instructions that have modified the value of eax and use any appropriate operands involved with the inverse operation of the one originally performed on the value to reverse the transformations to eax's value. Finally, we notice eax originally received taint when a value was moved from memory, which was read into memory using the read syscall.

FROM INTRO
We have implemented this technique in a proof of concept fuzzer, stubble, which operates on x86 binaries. Stubble dynamically instruments binaries, and automatically generates test files which exercise new paths in the input program. Program input, output, and interesting conditions like segfaults are recorded during execution, and taint propagation is recorded from user input sources to branch sinks. When a tainted branch is found, stubble automically backsolves the execution slice and generates a new test case which will drive the program down the opposite path explored in the current execution. Stubble is able to reason about large amounts of bytes of user input, for instance by performing oracle attacks on binaries which check user input against large secrets, requiring O(n) test cases to solve the secret with no previous information, where n is the length of the secret. Stubble has also been parallized at the process level, allowing for an arbitrary number of independent bytes to be reasoned about concurrently. We believe automated dynamic slice backsolving is a reasonable complementary technique to dynamic taint analysis and symbolic execution for binary fuzzing.
END FROM INTRO

@@

Stubble architecture, describe
why intel pin and not dynamo rio, valgrind, etc.

@@

\section{Limitations}

@@

@@
